\section{Modularizing Handshake Encryption}
\label{sec:modularizing}
%\TODO{possibly rename $KE_1$ and $KE_2$ in section 5 to match naming in corollary section}

Next will argue that using ``internal'' keys to encrypt handshake messages on the TLS~1.3 record-layer does not impact the security of other keys established by the handshake. 
%This will then enable us to use an as-simple-as-possible key exchange security model.
%In particular, we avoid considering such record-layer encryption within the model, as in ACCE-style security models~\cite{C:JKSS12}, for instance, which reduces the complexity of the model and proofs significantly. 
%

	Theorem~\ref{thm:maul}
		below %
formulates our argument in a general way, applicable to any multi-stage key exchange protocol, so that future analyses of similar protocols might take advantage of this modularity as well.

Intuitively, we argue as follows. Let $\KE_{2}$ be a protocol that provides multiple different stages with different external keys (i.e., none of the keys is used in the protocol, e.g., to encrypt messages), and let $\KE_{1}$ be the same protocol, except that some keys are ``internal'' and used, e.g., to encrypt certain protocol messages.
We argue that either using ``internal'' keys in $\KE_{1}$ does not harm the security of \emph{other} keys of $\KE_{1}$, or $\KE_{2}$ cannot be secure in the first place.
This will establish that we can prove security of a variant TLS~1.3 \emph{without} handshake encryption (in an accordingly simpler model),
and then lift this result to the actual TLS~1.3 protocol \emph{with} handshake encryption and the handshake traffic keys treated as ``internal'' keys.


\begin{theorem}
	\label{thm:TLS-transform}
	Let $\KE_1$ be the TLS~1.3 PSK-only resp.\ PSK-(EC)DHE mode \emph{with} handshake encryption (i.e., with internal stages $\KE_1.\INT = \{3,4\}$) as specified on the right-hand side in Figure~\ref{fig:tls-handshake}.
	Let $\KE_2$ be the same mode \emph{without} handshake encryption (i.e., $\KE_1.\INT = \emptyset$ and AEAD-encryption/decryption of messages is omitted).
	Let $\MaulSend$ and $\MaulRecv$ be the AEAD encryption resp.\ decryption algorithms deployed in TLS~1.3 and $\MaulKeys = \KE_1.\INT = \{3,4\}$.
Then we have	
%	For any adversary~$\advA$ against the multi-stage key exchange security of~$\KE_2$, running in time $t$ and making $\qSend$ queries to its $\Send$ oracle,
%	there exists an adversary~$\advB$ that runs in time $\approx t + \qSend \cdot t_{\mathrm{AEAD}}$ (where $t_{\mathrm{AEAD}}$ is the maximum time required to execute AEAD encryption or decryption of TLS~1.3 messages) and makes at most $\qSend$ queries to $\RevSessionKey$ in addition to queries made by $\advA$, and the same number of queries as~$\advA$ to all other oracles in the $\KESEC$ game,
%	such that
	\shortlongeqn[,]{
		\Adv^{\KESEC}_{\KE_1}(t, \qNewSecret, \qSend, \qRevSessionKey, \qRevLongTermKey, \qTest, \qRO)
		\leq
		\Adv^{\KESEC}_{\KE_2}(t+ t_{\mathrm{AEAD}} \cdot \qSend, \qNewSecret, \qSend, \qRevSessionKey + \qSend, \qRevLongTermKey, \qTest, \qRO)
% 		\genAdv{\KESEC}{\KE_2}{\advA} \leq \genAdv{\KESEC}{\KE_1}{\advB}.
	}
	where $t_{\mathrm{AEAD}}$ is the maximum time required to execute AEAD encryption or decryption of TLS~1.3 messages.
\end{theorem}




% we establish the following formal result saying that to analyze security of the TLS~1.3 PSK-only or PSK-(EC)DHE mode \emph{with} handshake encryption (captured as~$\KE_2$), it suffices to analyze the security of the same protocol \emph{without} handshake encryption (denoted $\KE_1$)---in an accordingly simpler model only treating ``external'' keys---, on which we can apply the AEAD encryption/decryption transformation.


For TLS~1.3 this means that we will not consider any security guarantees provided by the additional encryption of handshake messages.
We consider this as reasonable for PSK-mode ciphersuites, because the main purposes of handshake message encryption in TLS~1.3 is to hide the identities of communicating parties, e.g., in digital certificates, cf.\ \cite{PoPETS:ABFNO19}. 
In PSK mode there are no such identities. The $\pskid$ might be viewed as a string that could identify communicating parties, but it is sent unencrypted in the $\ClientHello$ message, anyway; the encryption of subsequent handshake messages would not contribute to its protection. 
%Therefore we consider it reasonable to essentially ignore the additional privacy-protecting properties of handshake message encryption in our model and proofs.


% For the case of the TLS~1.3 PSK modes,
% we will treat as $\KE_2$ the TLS~1.3 PSK-only resp.\ PSK-(EC)DHE mode as specified on the left-hand side of Figure~\ref{fig:tls-handshake} (i.e., \emph{with} handshake encryption using the internal keys from stages~3 and~4 ($\chtk$ and $\shtk$),
% and as $\KE_1$ the corresponding PSK handshake \emph{without} handshake encryption.


%\TODO{update Theorem 3 in Section 5 to take place after Section 4}

\iffull
	\input{tls-thesis/appendix-modular-transform}
\else
\fi

 
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:
