\section{Notes}

Notes from meeting 9/10/2020: 
\begin{itemize}
\item tight PSK/resumption mode  [here the PSK would be the long-term secret]
\begin{itemize}
\item does some kind of mu-PRF security with Corrupts work?
\item can we achieve tightness when modeling HKDF.Extract/Expand as RO,
implicitly programming the RO when we give out a PSK upon Corrupt query?
\end{itemize}
\item tight PSK + EC-DHE mixed

\item tight 0-RTT mode  (somewhat part of the PSK modes above)
\begin{itemize}
\item needs multi-user session resumption?

\end{itemize}
\item Backwards compatibility/downgrade resilience vs. tightness
\begin{itemize}
	\item prior TLS~1.3 results rely on hash CR + signatures; is there
anything to gain in terms of tightness?
\end{itemize}

\item Can we generalize our results/techniques? What's the type of key
exchange amenable to this?
\begin{itemize}
	\item  if all we know is SIGMA or HMQV-like, there's only little gain?
\end{itemize}

\item Is there a more modular appraoch via, e.g., a multi-user strong-DH
assumption?

\item Memory tightness
\begin{itemize}
	\item currently, we're linear in \#sessions/RO-queries?
\item DD/TJ ongoing work; best to wait for these
results first and see how they could be applied
\end{itemize} 
\item Tight signatures are a remaining bottelneck
\begin{itemize}
	\item  ongoing work with Eike Kilitz
\end{itemize}

\item Signal initial key exchange (X3DH)

\begin{itemize}
	\item can this be shown tightly BR-secure using our technique?
(of course, Signal actually wants eCK security, but might still be
interesting)
\end{itemize}
\item Tight eCK security
\begin{itemize}
\item possible/impossible? ... that's a tough one :-)
\end{itemize}
\item LWE-based key exchange / NewHope
\begin{itemize}
\item NewHope NIST submission is a KEM, can be used for ephemeral KEM +
SIGMA-style authentication (cf. https://ia.cr/2017/1252)
\item can we use some self-reducibility of LWE?
\end{itemize}
\end{itemize}

Additional notes from meeting 9/24/2020
\begin{itemize}
	\item Best approach to start with PSK \& 0-RTT modes, then extend further.
	\item cut down PSK MSKE model, add RO modeling and explicit authentication
	\item FG+HD RO modeling should be compatible with MSKE model, assuming only final-stage keys (SFK, CFK, tkshs,tkchs, ATS) are modeled as intermediate keys. 
	\item Keep a close eye on symmetric numbers: in ACNS draft, a loose bound on collisions dominated when using large curves. Might be a problem for PSK w/no curves. 
	\item LWE-based key exchange is likely not suited to this kind of approach because the gap-LWE problem is not hard.
	\item Look at KEM-based AKE protocols such as https://eprint.iacr.org/2020/1088.pdf
\end{itemize}