\section{Tight Security of the TLS~1.3 PSK Modes}
\label{sec:ke-proof}

In this section, we apply the insights gained in Sections~\ref{sec:ks-indiff} and~\ref{sec:modularizing} to obtain tight security bounds for both the PSK-only and the PSK-(EC)DHE mode of TLS~1.3.
To that end, we first present the protocol-specific properties of the TLS~1.3 PSK-only and PSK-(EC)DHE modes such that they can be viewed as multi-stage key exchange (MSKE) protocols as defined in Section~\ref{sec:ake-model}.
Then, we prove tight security bounds in the MSKE model in Theorem~\ref{thm:full-psk-ecdhe-ke} for the TLS~1.3 PSK-(EC)DHE mode and 
	in Theorem~\ref{thm:psk-ke} 
for the TLS~1.3 PSK-only mode%
	.

\subsection{TLS~1.3 PSK-only/PSK-(EC)DHE as a MSKE Protocol}
\label{sec:tls-formal-def}

We begin by capturing the TLS~1.3 PSK-only and PSK-(EC)DHE modes, specified in \autoref{fig:tls-handshake}, formally as MSKE protocols.
To this end, we must explicitly define the variables discussed in Section~\ref{sec:ake-model}.
In particular, we have to define the stages themselves, which stages are internal and which replayable, the session and contributive identifiers, when stages receive explicit authentication, and when stages become forward secret.

\paragraph{Stages.}
The TLS~1.3 PSK-only/PSK-(EC)DHE handshake protocol has eight stages (i.e., $\STAGES = 8$), corresponding to the keys $\ets$, $\eems$, $\shtk$, $\chtk$, $\cats$, $\sats$, $\ems$, and $\rms$ in that order.
The set $\INT$ of internal keys contains $\chtk$ and $\shtk$, the handshake traffic encryption keys. 
Stages $\ets$ and $\eems$ are replayable: $\REPLAY[s]$ is true for $s \in \{1,2\}$ and false for all others.

\paragraph{Session and contributive identifiers.}
The session and contributive identifiers for stage$s$ are tuples $(\mathit{label}_s, \mathit{ctxt})$, where $\mathit{label}_s$ is a unique label identifying stage~$s$, and $\mathit{ctxt}$ is the transcript that enters key's derivation.
The session identifiers $(\sid[s])_{s \in \{1, \dotsc, 8\}}$ are defined as follows:%
\footnote{Components marked with ${}^\dagger$ are only part of the TLS~1.3 PSK-(EC)DHE handshake.}
%
{\allowdisplaybreaks
\begin{align*}
	\iffull
	\sid[1] &= \left(``\ets", (\mathlist{\CH, \CKS^\dagger, \CPSKtls}) \right), \\
	\sid[2] &= \left( ``\eems", (\mathlist{\CH, \CKS^\dagger, \CPSKtls}) \right), \\
	\else
	\sid[1] / \sid[2] &= \left(``\ets" / ``\eems", (\mathlist{\CH, \CKS^\dagger, \CPSKtls}) \right), \\
	\fi
	\iffull
	\sid[3] &= \left( ``\chtk", (\mathlist{\CH, \CKS^\dagger, \CPSKtls, \SH, \SKS^\dagger, \SPSKtls}) \right), \\
	\sid[4] &= \left( ``\shtk", (\mathlist{\CH, \CKS^\dagger, \CPSKtls, \SH, \SKS^\dagger, \SPSKtls}) \right), \\
	\else
	\sid[3] / \sid[4] &= \left( ``\chtk" / ``\shtk", (\mathlist{\CH, \CKS^\dagger, \CPSKtls, \SH, \SKS^\dagger, \SPSKtls}) \right), \\
	\fi
	\iffull
	\sid[5] &= \left( ``\cats", (\mathlist{\CH, \CKS^\dagger, \CPSKtls, \SH, \SKS^\dagger, \SPSKtls, \EE, \SF}) \right), \\
	\sid[6] &= \left( ``\sats", (\mathlist{\CH, \CKS^\dagger, \CPSKtls, \SH, \SKS^\dagger, \SPSKtls, \EE, \SF}) \right), \\
	\sid[7] &= \left( ``\ems", (\mathlist{\CH, \CKS^\dagger, \CPSKtls, \SH, \SKS^\dagger, \SPSKtls, \EE, \SF}) \right), \text{ and} \\
	\else
	\sid[5] / \sid[6] / \sid[7] &= \left( ``\cats" / ``\sats" / ``\ems", (\mathlist{\CH, \dots, \SPSKtls, \EE, \SF}) \right), \text{ and} \\
	\fi
	\sid[8] &= \left( ``\rms", (\mathlist{\CH, \CKS^\dagger, \CPSKtls, \SH, \SKS^\dagger, \SPSKtls, \EE, \SF, \CF}) \right).
\end{align*}}%
%
To make sure that a server that received $\ClientHello$, $\ClientKeyShare^\dagger$, and $\ClientPreSharedKey$ untampered can be tested in stages~$3$ and~$4$,
even if the sending client did not receive the server's answer,
we set the contributive identifiers of stages~$3$ and~$4$ such that~$\rolecid$ reflects the messages that a session in role~$\role$ must have honestly received for testing to be allowed.
Namely, we let clients (resp. servers) upon sending (resp. receving) the messages $(\CH, \CKS^\dagger, \CPSKtls)$ set
\iffull
\begin{align*}
	\respcid[3] &= \left( ``\chtk", (\mathlist{\CH, \CKS^\dagger, \CPSKtls}) \right) \text{ and}\\
	\respcid[4] &= \left( ``\shtk", (\mathlist{\CH, \CKS^\dagger, \CPSKtls}) \right).
\end{align*}
\else
$\respcid[3] = (``\chtk", (\mathlist{\CH, \CKS^\dagger, \CPSKtls}))$ and $\respcid[4] = (``\shtk", (\mathlist{\CH, \CKS^\dagger, \CPSKtls}))$.
\fi
Further, when the client receives (resp. the server sends) the message $(\mathlist{\SH, \SKS^\dagger, \SPSKtls})$, they set
\iffull
\[
	\initcid[3] = \sid[3] \quad\text{and}\quad \initcid[4] = \sid[4].
\]
\else
$\initcid[3] = \sid[3]$ and $\initcid[4] = \sid[4]$.
\fi
For all other stages $s \in \{\mathlist{1,2,5,6,7,8}\}$, $\initcid[s] = \respcid[s] = \sid[s]$ is set upon acceptance of the respective stage (i.e., when $\sid[s]$ is set as well).


\paragraph{Explicit authentication.}
For initiator sessions, all stages achieve explicit authentication when the $\ServerFinished$ message is verified successfully.
This happens right before stage~$5$ (i.e., $\cats$) is accepted.
That is, upon accepting stage~$5$ all previous stages receive explicit authentication retroactively and all following stages are explicitly authenticated upon acceptance.
Formally, we set $\EAUTH[\initiator, s] = 5$ for all stages $s\in \{1,\dotsc,8\}$.

\iffull
For responder session, all stages receive explicit authentication upon (successful) verification of the $\ClientFinished$ message. 
This occurs right before the acceptance of stage~$8$ (i.e., $\rms$).
Similar to initiators, responders receive explicit authentication for all stages upon acceptance of stage~$8$ since this is the last stage of the protocol.
Accordingly, we set $\EAUTH[\responder, s] = 8$ for all stages $s \in \{1,\dotsc,8\}$.
\else
Analogously, responder sessions receive explicit authentication right before accepting stage~$8$ via the $\ClientFinished$ message;
i.e., $\EAUTH[\responder, s] = 8$ for all stages $s \in \{1,\dotsc,8\}$.
\fi

\paragraph{Forward secrecy.}
Only keys dependent on a Diffie--Hellman secret achieve forward secrecy, so all stages~$s$ of the PSK-only handshake have $\FS[r, s, \fs] = \FS[r, s, \wfstwo] = \infty$ for both roles~$r \in \{\initiator, \responder\}$.
In the PSK-(EC)DHE handshake, full forward secrecy is achieved at the same stage as explicit authentication for all keys except $\ets$ and $\eems$, which are never forward secret.
That is, for both roles~$r$ and stages $s \in \{3,\dotsc,8\}$ we have $\FS[r, s, \fs] = \EAUTH[r, s]$.
All keys except $\ets$ and $\eems$ possess weak forward secrecy $2$ upon acceptance, so we set $\FS[r, s, \wfstwo] = s$ for stages~$s \in \{3,\dotsc,8\}$.
Finally, as stages~$1$ and~$2$ (i.e., $\ets$ and $\eems$) never achieve forward secrecy we set $\FS[r, s, \fs] = \FS[r, s, \wfstwo] = \infty$ for both roles~$r$ and stages $s \in \{1,2\}$.

% paragraph is in the intro now
%\paragraph{\replace{View on the TLS~1.3 handshake protocol}{Scope of our analysis}.}
%Like many previous cryptographic analyses~\cite{C:JKSS12,C:KraPatWee13,CCS:DFGS15,EuroSP:FisGue17,JC:DFGS21,JC:DieJag21,ACNS:DavGun21} of the TLS handshake protocol, we focus only on the ``cryptographic core'' of the PSK modes of the TLS~1.3 handshake protocol.
%In particular, our analysis given in Section~\ref{sec:psk-ecdhe-ke-proof} (resp.\ Section~\ref{sec:psk-only-ke-proof}) focuses solely on the TLS~1.3 PSK-(EC)DHE (resp.\ TLS~1.3 PSK-only) handshake protocol (with optional $0$-RTT) for a fixed Diffie--Hellman group and for a fixed cipher suite (i.e., AEAD algorithm and hash algorithm).
%This means we neither consider the negotiation of other versions of TLS than version 1.3 nor consider the negotiation of the DH group for the PSK-(EC)DHE mode and the cipher suite for both PSK-only and PSK-(EC)DHE mode,
%as done, e.g., in~\cite{ACISP:DowSte15,SP:BBFGKZ16}.
%Since we only consider TLS~1.3 ``in isolation'' we also do not consider backward compatibility.
%\dd{This was a big question by one reviewer on our JoC submission.}
%Moreover, our view does not include the TLS record layer
%and we we do not consider the composition of the handshake protocol with the record layer encryption as, e.g., done in \cite{CCS:DFGS15,thesis:Guenther18,JC:DFGS21,JC:DieJag21}.
%
%\TODO{find a better caption.}
%\dd{Pls check if anything important is missing (or if the record layer part seems misplaced.)}

\subsection{Tight Security Analysis of TLS~1.3 PSK-(EC)DHE}
\label{sec:psk-ecdhe-ke-proof-short}

We now come to the tight MSKE security result for the TLS~1.3 PSK-(EC)DHE handshake.

\begin{theorem}
	\label{thm:full-psk-ecdhe-ke}
	Let $\TLSPSKECDHE$ be the TLS~1.3 PSK-(EC)DHE handshake protocol (with optional 0-RTT) as specified on the right-hand side in \autoref{fig:tls-handshake} without handshake encryption.
	Let $\G$ be the Diffie--Hellman group of order~$p$.
	Let $\nl$ be the length in bits of the nonce, let $\hashlen$ be the output length in bits of $\abstractHash$, and let the pre-shared key space be $\KEpskeyspace = \bits^\hashlen$.
	We model the functions $\abstractHash$ and $\TLSKDF_x$ for each $x \in \{\binder, \dotsc, \rms\}$ as $12$ independent random oracles $\mathlist{\RO_\Thash,\RO_{\binder},\dotsc, \RO_{\rms}}$.
	%
% 	Let $\advA$ be an adversary against the $\KESEC$ security of $\TLSPSKECDHE$ running in time $t$, and let $\advA$ make $\qNewSecret$, resp.\ $\qSend$ queries to its $\NewSecret$, resp.\ $\Send$ oracles.
% 	Further, let $\qRO$ be the number of random oracle queries $\advA$ makes in total to its random oracles $\mathlist{\RO_\Thash,\RO_{\binder},\dotsc, \RO_{\rms}}$.
	Then,
	\begin{align*}
		&\genAdv{\KESEC}{\TLSPSKECDHE}{t, \qNewSecret, \qSend, \qRevSessionKey, \qRevLongTermKey, \qTest, \qRO}
		\leq \frac{2\qSend^2}{2^{\nl} \cdot p} \\
		&+ \frac{(\qRO+\qSend)^2 + \qNewSecret^2 + (\qRO+6\qSend)^2 + \qRO \cdot \qNewSecret + \qSend}{2^{\hashlen}}
		+  \frac{4(t + 4 \log(p) \cdot \qRO)^2}{p}.
	\end{align*}
\end{theorem}

\begin{remark}
	Our MSKE model from Section~\ref{sec:ake-model} assumes pre-shared keys to be uniformly random sampled from~$\KEpskeyspace$, where here $\KEpskeyspace = \bits^\hashlen$.
	This matches how pre-shared keys are derived for session resumption, as well as our analysis of domain separation, which assumes pre-shared keys to be of length~$\hashlen$.
\end{remark}

\begin{remark}\label{rem:psk-distribution}
	Our bound is easily adapted to any distribution on $\bits^{\hashlen}$ in order to accommodate out-of-band pre-shared keys that satisfy the length requirement but do not have full entropy.
	%Since we need to avoid collisions for the pre-shared keys, we obtain lower bound on the distributions entropy for a desired security level.
	Expectedly, lower-entropy PSK distributions result in weaker bounds, due to the increased chance for collisions between PSKs as well as the adversary guessing a PSK.
\end{remark}

\iffull
	\begin{proof}
	\input{tls-thesis/full-psk-dhe-proof}
	\end{proof}
\else
\subsection{Proof overview}
\label{sec:psk-ecdhe-ke-proof-overview}

The proof proceeds via a sequence of games in three phases, corresponding to the three ways for an adversary to win the $\KESEC$ security game.
We begin with $\Gm_0$, the original $\KESEC$ game for protocol $\TLSPSKECDHE$ described above. 
In the first phase, we establish that the adversary cannot violate the $\Sound$ predicate.
In the second phase, we establish the same for the $\ExplicitAuth$ predicate.
In the third phase, we ensure that all $\Test$ queries return random keys regardless of the value of the challenge bit~$b$, so long as the $\Fresh$ predicate is not violated.
After that, the adversary cannot win the game with probability better than guessing, rendering its advantage to be~$0$.
We bound the advantage difference introduced by each game hop; collecting these intermediate bounds yields the overall bound.
For space reasons, we only provide a summary of the proof in the following and refer to 
	Appendix~\ref{app:full-psk-ecdhe-ke-proof} 
for the full details.


\subsection*{Phase~1: Ensuring $\Sound$}
The $\Sound$ predicate 
	(cf.\ Figure~\ref{fig:MSKE-preds}) 
checks that no more than two sessions can be partnered in a non-replayable stage, and that any two partnered sessions must agree on the stage, pre-shared key identifier, the stage-$s$ key, and each others' identities and roles. 
We defined our session identifiers so that the stage-$s$ session identifier contains (1) a label unique to that stage,  (2) a unique $\ClientHello$ and $\ServerHello$ message, (3) the $\binder$ message: a $\MAC$ tag authenticating the $\ClientHello$ and pre-shared key, and (4) sufficient information to fix the stage-$s$ key. 
(This does not mean the key is computable from the $\sid$; it is not.)

We then perform three incremental game hops that cause the $\Finalize$ oracle to return $0$ in the event of a collision between two $\Hello$ messages, $\binder$ tags, or pre-shared keys.
We bound the difference in advantage in the first two game hops via a birthday bound over the number of potentially colliding values (i.e., pairs of nonces and $\KeyShares$ in $\G$ for $\Hello$ message collisions, and sampled $\psk$ keys for pre-shared key collisions),
and the third hop by a reduction to the collision resistance of the $\RObinder$ random oracle whose advantage in turn is upper bounded by a birthday bound~$\genAdv{\CollRes}{\RObinder}{\qRO + \qSend} \leq \frac{(\qRO + \qSend)^2}{2^\hashlen}$%
	; cf.\ Appendix~\ref{app:coll-res-ro}.
The resulting bounds are, in this order:
\shortlongeqn[.]{
	\Pr[\Gm_0] - \Pr[\Gm_3] \leq \frac{2\qSend^2}{2^{\nl} \cdot p} + \frac{\qNewSecret^2}{2^{\hashlen}} + \frac{(\qRO + \qSend)^2}{2^\hashlen}
}
As long as no such collisions occur, each stage-$s$ session identifier uniquely determines one client session, one server session (for non-replayable stages), one pre-shared key (and therefore one peer and identifier owning that key), and one stage-$s$ session key.
At this point, the $\Sound$ predicate will always be $\true$ unless $\Finalize$ would return $0$, so the adversary cannot win by violating $\Sound$.
%\fg{This is a summary for what would be a combined proof for PSK-only and PSK-(EC)DHE; separate.}
%\replace{In the PSK-(EC)DHE handshake, h}{H}

\subsection*{Phase~2: Ensuring $\ExplicitAuth$}
In the second phase of the proof, we change the key-derivation process to avoid sampling pre-shared keys wherever possible,
instead replacing keys and $\MAC$ tags derived from those pre-shared key by uniformly random strings.
We then make the adversary lose if it makes queries that would allow him to detect these changes and bound that probability;
in particular we ensure that the adversary does not correctly guess a now-random $\ClientFinished$ or $\ServerFinished$ MAC tag.
Sessions achieve explicit authentication just after verifying their received $\Finished$ message;
eliminating possible forgeries hence ensures that the $\ExplicitAuth$ predicate cannot be $\false$ without $\Finalize$ returning $0$.
All changes in this phase apply only to sessions whose pre-shared key has not been corrupted.

\lightparagraph{Game~4}
Our first of six game hops eliminates collisions in the ``transcript hash'' function $\ROthash$.
We reduce to the collision resistance of $\ROthash$ and bound this advantage with a birthday bound:
\shortlongeqn[.]{
	\Pr[\Gm_3] - \Pr[\Gm_4] \leq \frac{\qRO+6\qSend}{2^{\hashlen}}
}
(The factor~$6$ comes from the up to~$6$ transcript hashes computed in any~$\Send$ query.)

\lightparagraph{Game~5}
Our next game forces $\Finalize$ to return $0$ if the adversary guesses any uncorrupted pre-shared key in any random oracle query.
Since we assume pre-shared keys are uniformly random,
\shortlongeqn[.]{
	\Pr[\Gm_4] - \Pr[\Gm_5] \leq \frac{\qRO \cdot \qNewSecret}{2^{\hashlen}}
}

\lightparagraph{Games~6 and~7}
In our third game hop, we ask log the stage~$s$ key computed in any session in a look-up table~$\skeyTable$ under its session identifier.
Sessions whose partners have logged a key can then, in a fourth game hop, copy the key from $\skeyTable$ instead of deriving it.
Partnered sessions will always derive the same key as guaranteed by the $\Sound$ predicate, so the adversary cannot detect the copying and its advantage does not change.
In addition to logging and copying keys, we also log and copy the three $\MAC$ tags: $\binder$, $\sfin$, and $\cfin$ using another look-up table~$\macTable$. 
Since $\MAC$ tags do not have associated session identifiers, they are logged under the inputs to~$\RObinder$, $\ROsfin$, resp.\ $\ROcfin$.
This technique is inspired by the work of Cohn-Gordon et al.~\cite{C:CCGJJ19}. 

\lightparagraph{Game~8}
In preparation for the final step in this phase, our fifth game hop eliminates uncorrupted pre-shared keys altogether.
We postpone the sampling of the pre-shared key to the $\RevLongTermKey$ oracle so that only corrupted sessions hold pre-shared keys.
As a consequence of this change, we can no longer compute session keys and MAC tags using the random oracles.
Sessions will instead sample these uniformly at random from their respective range. 
In another look-up table, they log the $\RO$ queries they would have made so that these queries can be programmed later if the pre-shared key gets corrupted.
Queries to~$\RO$ before corruption cannot contain the pre-shared key thanks to the previous game, so we do not have to worry about consistency with past queries. 
%
We also cannot  implement the previous games' check for guessed pre-shared keys in $\RO$ queries until these keys are sampled,
so we sample new pre-shared keys for all uncorrupted identifiers at the end of the game in the $\Finalize$ oracle, then perform the check.
The programming of the random oracles is perfectly consistent with their responses in earlier games, so the adversary cannot detect when pre-shared keys are chosen in the game and its advantage does not change.
%We don't need to 

\lightparagraph{Game~9}
The final game in this phase ensures that either $\ExplicitAuth = \true$ or $\Finalize$ returns~$0$.
In this game, we return $0$ from $\Finalize$ if any honest session would accept the first explicitly-authenticated stage (stage~$5$ ($\cats$) for initiators and stage~$8$ ($\rms$) for responders) with an uncorrupted pre-shared key and no honest partner.
By the previous game, we established that sessions with uncorrupted pre-shared keys randomly sample their $\MAC$ tags,
unless they copy a cached result in which case the same computation was made by another session.
Thanks to the way we defined our session identifiers, no unpartnered session will copy their $\MAC$ tags:
the computation of the $\ServerFinished$ $\MAC$ tag contains the hash of the stage-$5$ $\sid$ (excluding $\sfin$); likewise the $\ClientFinished$ tag contains the hash of the stage-$8$ $\sid$.
Since we ruled out hash collisions in the first game of the phase, any two sessions computing the same $\ServerFinished$ message are stage-$5$ partners and any two sessions computing the same $\ClientFinished$ message are stage-$8$ partners.
So any unpartnered session with an uncorrupted pre-shared key has a random $\MAC$ tag, and the odds of the adversary guessing such a tag is bounded by $\frac{\qSend}{2^{\hashlen}}$.
With the prior two games not changing the adversary's advantange, we have
\shortlongeqn[.]{
	\Pr[\Gm_5] - \Pr[\Gm_9] \leq \frac{\qSend}{2^{\hashlen}}
}

We are now guaranteed that any session accepting the stage that achieves explicit authentication without a corrupted pre-shared key has a partner in that stage.
The $\Sound$ predicate guarantees that the partner agrees on the peer and pre-shared key identities, which is sufficient to guarantee explicit authentication for all responder sessions.
For initiator sessions, we must also note that a partner in stage $5$ will become, upon their acceptance, a partner in stages~$6$ ($\sats$) and~$7$ ($\ems$), whose $\sid$s are identical to that of stage $5$ apart from their labels.
An initiator's stage-$5$ partner will only accept a $\ClientFinished$ message identical to the one sent by the initiator, at which point they will become a partner also in stage~$8$.
This ensures that the $\ExplicitAuth$ predicate can never be false unless one of the flags introduced in this phase causes $\Finalize$ to return~$0$.


\subsection*{Phase~3: Ensuring the Challenge Bit is Random and Independent}

Our goal in the third and last phase is to ensure that all session keys targeted by a $\Test$ query are uniformly random and independent of the challenge bit $b$ whenever the $\Fresh$ predicate is true.
Freshness ensures that no session key can be tested twice or tested and revealed in the same stage either by targeting the same session twice or two partnered sessions.
It also handles our three levels of forward secrecy.

We can already establish this for $\Test$ queries to sessions in non-forward secret stages~$1$ ($\ets$) and~$2$ ($\eems$).
These queries violate $\Fresh$ unless the sessions' pre-shared keys are never corrupted. 
Since $\Gm_8$, all sessions with uncorrupted pre-shared keys either randomly sample their session keys, or copy random keys from a partner session.
If one of these session keys is tested, it cannot have been output by another $\Test$ or $\RevSessionKey$ query without violating $\Fresh$.
Therefore the response to the $\Test$ query is a uniformly random string, independent of all other oracle responses and the challenge bit $b$.

The remaining stages ($3$--$8$) have weak forward secrecy $2$ until explicit authentication is achieved, then they have full forward secrecy.
These stages' keys may be tested even if the session's pre-shared key has been corrupted, so long as there is a contributive partner (or, in the case of full forward secrecy, that the corruption occurred after forward secrecy was achieved).
We use one last game hop to ensure these keys are uniformly random when they are tested.

\lightparagraph{Game~10}
In $\Gm_{10}$, we cause the $\Finalize$ oracle to return $0$ if the adversary should ever make a random oracle query containing the Diffie--Hellman secret $\dhe$ of an honest partnered session whose pre-shared key was corrupted.
Without such a query, all keys derived from a Diffie--Hellman secret sampled uniformly at random by the random oracles.

We bound the probability of this event via a reduction $\advB_\dhe$ to the strong Diffie--Hellman problem in group $\G$.
(Recall that $\G$ has order $p$ and generator~$g$.)
In this problem, the adversary $\advB_\dhe$ gets as input a strong DH challenge $(A = g^a, B = g^b)$ as well as access to an oracle $\stDH_a$ for the decisional Diffie--Hellman (DDH) problem with the first argument fixed. 
Given inputs $C \assign g^c$ and $W$ for any $c \in \Z_p$, $\stDH_a(C, W)$ returns $\true$ if and only if $W = g^{ac} = C^a$.
The goal of~$\advB_\dhe$ is to submit $Z$ to its $\Finalize$ oracle such that $Z = g^{ab}$.

The reduction $\advB_\dhe$ simulates $\Gm_{10}$ for the $\KESEC$ adversary $\advA$.
At a high level, it uses rerandomization to embed its strong DH challenge $A$, resp. $B$, into the key shares of every initiator session, resp. every partnered responder session.
To embed a challenge $A$ in its key share, a session samples a ``randomizer'' $\tau \sample \Z_p$, and sets its key share to $X \assign A \cdot g^{\tau}$. 
If $\advA$ should make an $\RO$ query containing the Diffie--Hellman secret associated with two embedded key shares, the reduction can detect this query with its DDH oracle.
It then extracts the solution to its strong DH challenge from the query's DH secret, calls the $\Finalize$ oracle, and wins its own game.

There are a few subtleties to the reduction, which requires us to extend the technique of CCGJJ~\cite{C:CCGJJ19}.
Unlike honest executions of the protocol, the reduction's simulated sessions with embedded key shares do not know their own secret Diffie--Hellman exponents.
If their pre-shared keys are never corrupted, this does not matter because session keys and $\MAC$ tags are randomly sampled.
Corrupted sessions, however, cannot use the random oracles to compute these values as they would in $\Gm_{10}$.
%
Instead, $\advB_\dhe$ samples session keys and $\MAC$ tags uniformly at random and uses several look-up tables to program random oracle queries and maintain consistency between sessions.

\iffull
Recall that since $\Gm_6$, partnered sessions store their random session keys in a table $\skeyTable$ under their session identifiers.
Since session identifiers do not contain Diffie--Hellman secrets, simulated sessions with embedded key shares can also log and copy their keys using $\skeyTable$, even if their pre-shared key has been corrupted.
Partnered sessions with embedded key shares need not maintain consistency with later random oracle queries, as such query will enable our reduction to call $\Finalize$ and immediately end the game before a response is required.

Logging $\MAC$ tags in the table~$\macTable$ however does require the Diffie--Hellman secret, so partnered sessions with embedded key shares must use a separate table $\rndTable$ to achieve consistency.
This table also logs the information used to compute a $\MAC$ tag, but instead of a Diffie--Hellman secret it stores the initiator and responder's randomizers.
It also stores the full transcript instead of its hash, so that the reduction can efficiently recall the initiator and responders' key shares.
\fi

With this infrastructure in place, the reduction proceeds in the following way.
Whenever a partnered session with embedded key share would need its Diffie--Hellman secret, it searches all past $\RO$ queries for this secret.
It looks up the initiator's stored randomizer $\tau$ and the responder's randomizer $\tau'$.
Then for each guess $\Z$ in a past $\RO$ query, the reduction queries the strong Diffie--Hellman oracle on the responder's key share $\SKS$ and $C \assign \Z \cdot g^{-\tau} $.
This query will return $\true$ if the adversary correctly guessed the Diffie--Hellman secret; in this case the reduction calls $\Finalize(Z \cdot g^{-\tau} \cdot g^{-\tau'})$ and solves its strong DH challenge.
Unpartnered sessions do the same thing, except that the responder has no randomizer; in response to the strong DH oracle answering $\true$ they hence merely program their session keys instead of calling $\Finalize$.
We emphasize that for tightness, it is crucial to maintain efficiency during this process.
We do so by only checking $\RO$ queries whose context matches the hashed protocol transcript; this ensures $\advB_\dhe$ makes at most $2$ $\stDH_a$ queries for each $\RO$ query.

After a session chooses its session key or $\MAC$ tag, it stores the chosen value, its transcript, and all known randomizers in a table $\rndTable$.
When the reduction answers future $\RO$ queries, it will use this table to check if a query contains the Diffie--Hellman secret of an accepted session using the strong DH oracle as above; if so, they program or call $\Finalize$ in the same way.

This reduction solves the strong Diffie--Hellman problem whenever the adversary makes an $\RO$ query containing a partnered session's Diffie--Hellman secret,
so for reduction $\advB_\dhe$ with runtime $t_{\advB_{\dhe}}$, we have
\shortlongeqn[.]{
	\Pr[\Gm_9] - \Pr[\Gm_{10}] \leq \genAdv{\stDH}{\G}{t_{\advB_\dhe}, 2\qRO}
}
Davis and G{\"u}nther gave a bound in the generic group model for the strong DH problem; applying their Theorem~3.3~\cite{ACNS:DavGun21} results in
\shortlongeqn[.]{
	\Pr[\Gm_9] - \Pr[\Gm_{10}] \leq \tfrac{t_{\advB_\dhe}^2}{p}
}
\smallskip


At this point in the proof, the adversary $\advA$ cannot possibly make a $\RO$ query that outputs any tested session key of a forward secret (full or wfs2) stage~$s$.
If the tested session's pre-shared key is uncorrupted, $\advA$ cannot make the query because of $\Gm_5$.
If the session has a contributive partner in stage~$s$, then from $\Gm_{10}$, $\advA$ cannot make the query because it contains the Diffie--Hellman secret of a partnered session.
If it has accepted with no contributive partner and a corrupted pre-shared key, then by the guarantees we established in Phase 2, the corruption must have occurred before forward secrecy and explicit authentication were achieved.

As a result, the output of any $\Test$ query (that does not violate $\Fresh$) is a random string, sampled by either a session or the $\RO$ oracle independently of all other game variables including the challenge bit $b$.
The adversary therefore has a probability no greater than $\frac{1}{2}$ of winning $\Gm_{10}$.
Collecting this probability with the other bounds between games in our sequence gives the proof. %\qed

\fi % end short proof overview

\subsection{Full Security Bound for TLS~1.3 PSK-(EC)DHE and PSK-only}
\label{sec:psk-ecdhe-ke-full-bound}

We can finally combine the results of Sections~\ref{sec:ks-indiff}, \ref{sec:modularizing}, and our key exchange bound above to produce fully concrete bounds for the TLS~1.3 PSK-(EC)DHE and PSK-only handshake protocols as specified on the left-hand side of Figure~$1$.
This bound applies to the protocol \emph{with handshake traffic encryption} and \emph{internal keys} when \emph{only modeling as random oracle}~$\ROhash$ the hash function~$\abstractHash$.

First, we define three variants of the TLS~1.3 PSK handshake:
\begin{itemize}
	\item $\KE_{0}$, as defined in Theorem~\ref{thm:full-ks-indiff} with handshake traffic encryption and one random oracle $\ROhash$.
		(This is the variant we want to obtain our overall result for.)	
		
	\item $\KE_{1}$, as defined in Theorem~\ref{thm:full-ks-indiff} with handshake traffic encryption and $12$ random oracles $\ROthash$, $\RObinder$, \dots, $\ROrms$.
	\item $\KE_{2}$: as defined in Theorem~\ref{thm:TLS-transform}, with no handshake traffic encryption and $12$ random oracles $\ROthash$, $\RObinder$, \dots, $\ROrms$.
\end{itemize}

Theorem~\ref{thm:full-ks-indiff} grants that
\shortlongalign[.]{
	\Adv^{\KESEC}_{\KE_{0}}(t, \qNewSecret, \qSend, \qRevSessionKey, \qRevLongTermKey, \qTest, \qRO)
	\leq
	\fullonly{&~}
	\Adv^{\KESEC}_{\KE_{1}}(t\cab \qNewSecret\cab \qSend\cab \qRevSessionKey\cab \qRevLongTermKey\cab \qTest\cab \qRO)
	\fullonly{\\&}
	+ \frac{2(12\qSend+\qRO)^2}{2^{\hashlen}}
	+ \frac{2\qRO^2}{2^{\hashlen}}
	+ \frac{8(\qRO+36\qSend)^2}{2^{\hashlen}}
}

Next, we apply Theorem~\ref{thm:TLS-transform}, yielding the bound 
	\shortlongeqn[,]{
	\Adv^{\KESEC}_{\KE_1}(t\cab \qNewSecret\cab \qSend\cab \qRevSessionKey\cab \qRevLongTermKey\cab \qTest\cab \qRO)
	\leq
	\Adv^{\KESEC}_{\KE_2}(t+ t_{\mathrm{AEAD}}\cdot \qSend\cab \qNewSecret\cab \qSend\cab \qRevSessionKey + \qSend\cab \qRevLongTermKey\cab \qTest\cab \qRO)
}
where $t_{\mathrm{AEAD}}$ is the maximum time required to execute AEAD encryption or decryption of TLS~1.3 messages. 

Theorem~\ref{thm:full-psk-ecdhe-ke} then finally and entirely bounds the advantage against the $\KESEC$ security of $\KE_{2}$.
\iffull
Collecting these bounds gives 
\begin{align*}
	\Adv^{\KESEC}_{\KE_{0}}(t, \qNewSecret, \qSend, \qRevSessionKey, \qRevLongTermKey, \qTest, \qRO)
	&\leq
	\Adv^{\KESEC}_{\KE_{1}}(t, \qNewSecret, \qSend, \qRevSessionKey, \qRevLongTermKey, \qTest, \qRO)
	\\&\quad
	+ \frac{2(12\qSend+\qRO)^2}{2^{\hashlen}}
	+ \frac{2\qRO^2}{2^{\hashlen}}
	+ \frac{8(\qRO+36\qSend)^2}{2^{\hashlen}}\\
	&\leq
	\Adv^{\KESEC}_{\KE_2}(t+ t_{\mathrm{AEAD}}\cdot \qSend, \qNewSecret, \qSend, \qRevSessionKey + \qSend, \qRevLongTermKey, \qTest, \qRO)
	\\&\quad
	+\frac{2(12\qSend+\qRO)^2 + 2\qRO^2 + 8(\qRO+36\qSend)^2}{2^{\hashlen}} \\
	&\leq \frac{2\qSend^2}{2^{\nl} \cdot p} + \frac{(\qRO+\qSend)^2 + \qNewSecret^2 + (\qRO+6\qSend)^2 + \qRO \cdot \qNewSecret + \qSend}{2^{\hashlen}} \\
	&\quad + \frac{4(t + t_{\mathrm{AEAD}}\cdot \qSend + 4 \log(p) \cdot \qRO)^2}{p} \\
	&\quad + \frac{2(12\qSend+\qRO)^2 + 2\qRO^2 + 8(\qRO+36\qSend)^2}{2^{\hashlen}}.
\end{align*}

This yields the following overall result for the $\KESEC$ security of the TLS~1.3 PSK-(EC)DHE handshake protocol.
\else
Collecting these bounds yields the following overall result for the $\KESEC$ security of the TLS~1.3 PSK-(EC)DHE handshake protocol.
\fi

\begin{corollary}\label{cor:full-psk-ecdhe-ke}
	Let $\TLSPSKECDHE$ be the TLS~1.3 PSK-(EC)DHE handshake protocol as specified on the left-hand side in \autoref{fig:tls-handshake}.
	Let $\G$ be the Diffie--Hellman group of order~$p$.
	Let $\nl$ be the length in bits of the nonce, let $\hashlen$ be the output length in bits of $\abstractHash$, and let the pre-shared key space be $\KEpskeyspace = \bits^\hashlen$.
	Let $\abstractHash$ be modeled as a random oracle $\RO_\hash$.
	%
% 	Let $\advA$ be an adversary against the $\KESEC$ security of $\TLSPSKECDHE$ running in time $t$, and let $\advA$ make $\qRO$, $\qNewSecret$, resp.\ $\qSend$ queries to its $\RO_\hash$, $\NewSecret$, resp.\ $\Send$ oracles.
	Then,
{\allowdisplaybreaks
	\begin{align*}
		&\genAdv{\KESEC}{\TLSPSKECDHE}{t, \qNewSecret, \qSend, \qRevSessionKey, \qRevLongTermKey, \qTest, \qRO} \\
		&\qquad\leq \frac{2\qSend^2}{2^{\nl} \cdot p} + \frac{(\qRO+\qSend)^2 + \qNewSecret^2 + (\qRO+6\qSend)^2 + \qRO \cdot \qNewSecret + \qSend}{2^{\hashlen}} \\
		&\qquad\qquad +  \frac{4(t + t_{\mathrm{AEAD}} \cdot \qSend + 4 \log(p) \cdot \qRO)^2}{p}\\
		&\qquad\qquad + \frac{2(12\qSend+\qRO)^2 + 2\qRO^2 + 8(\qRO+36\qSend)^2}{2^{\hashlen}}.
	\end{align*}
}
\end{corollary}

	\input{tls-thesis/psk-only-pf}


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:
