


\newif\ifacm
\newif\ifllncs
\newif\ifsubmission
\newif\ifanonymous
\newif\iffull
\newif\ifuseshortbib
\newif\ifdraft



\newif\iftikzEnabled
\tikzEnabledtrue

\useshortbibfalse
\acmfalse
\llncsfalse
\submissiontrue
\anonymousfalse
\fulltrue
%\drafttrue

\ifanonymous
\newcommand{\ourauthors}{}
\fi

\pagestyle{plain}

\ifacm \documentclass{sig-alternate}
\else \ifllncs  \documentclass[twoside,envcountsame,runningheads]{llncs}
\else \documentclass[11pt,twoside]{article}
         \usepackage[letterpaper,hmargin=1in,vmargin=1in]{geometry}
\fi\fi


\newcommand\hmmax{0} % tried to avoid too many fonts error
\newcommand\bmmax{0} % tried to avoid too many fonts error
\usepackage{amsmath}
\usepackage{enumerate}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{breakcites}
\usepackage{color}
\usepackage{float}
\usepackage{stmaryrd}
\usepackage{listings}
\usepackage{cryptocode}
\usepackage{crypto-environments}
\usepackage[normalem]{ulem}
\usepackage{tikz}
\usetikzlibrary{matrix,positioning,automata,shapes.geometric}
%\pgfdeclarelayer{background}
%\pgfsetlayers{background,main}

\ifanonymous
 \usepackage[
    pdftitle={},
    bookmarks=false,
    pdfauthor={}, pdfpagelabels=true, linktocpage=true]{hyperref}
\else
    \usepackage[
    pdftitle={},
    bookmarks=false,
    pdfauthor={Bellare, Davis, Di},
    pdfpagelabels=true, linktocpage=true, hyperfootnotes=false,
    pagebackref% references from biblio to pages where cited
    ]{hyperref}
\fi

\def\Bellare{Mihir Bellare}
\def\BellareEmail{\tt mbellare@ucsd.edu}
\def\BellareWebpage{\tt http://cseweb.ucsd.edu/\textasciitilde mihir/}
\def\BellareDept{Department of Computer Science \& Engineering}
\def\BellareCompany{University of California, San Diego}
\def\BellareAddress{9500 Gilman Drive, La Jolla, California 92093, USA}


\def\BellareThanks{
\BellareDept,
\BellareCompany,
\BellareAddress.
Email: {\tt \BellareEmail}.
URL: {\tt \BellareWebpage}.
Supported in part by NSF grant CNS-2154272.
}

\def\Davis{Hannah Davis}
\def\DavisEmail{\tt h3davis@ucsd.edu}
\def\DavisWebpage{\tt http://cseweb.ucsd.edu/\textasciitilde h3davis/}
\def\DavisDept{Department of Computer Science \& Engineering}
\def\DavisCompany{University of California, San Diego}
\def\DavisAddress{9500 Gilman Drive, La Jolla, California 92093, USA}

\def\DavisThanks{
	\DavisDept,
	\DavisCompany,
	\DavisAddress.
	Email: {\tt \DavisEmail}.
	URL: {\tt \DavisWebpage}.
	Supported in part by NSF grant CNS-2154272.
}


\ifllncs\else
\newcommand{\ourauthors}{
  \textsc{\Bellare}\thanks{\ \BellareThanks}
}
\fi

\input{header}

\newcommand{\titletext}{Hardening Signature Schemes via Derive-then-Derandomize: \\
Stronger Security Proofs for EdDSA}

\newcommand{\titletextnocr}{Hardening Signature Schemes via Derive-then-Derandomize: Stronger Security Proofs for EdDSA}

%Security Proofs for EdDSA Signatures with MD-hash functions
%Filtered Indifferentiability: \\  Security Proofs for EdDSA with SHA512

\newcommand{\shorttitletext}{Hardening Signature Schemes}
% Filtered Indifferentiability:  Security Proofs for EdDSA with SHA512

\ifllncs
\titlerunning{\shorttitletext}
\ifanonymous \authorrunning{} \else \authorrunning{Mihir Bellare, Hannah Davis, Zijing Di} \fi
\fi



\allowdisplaybreaks

\begin{document}

\setlength{\abovedisplayskip}{2pt}
\setlength{\belowdisplayskip}{2pt}
\setlength{\abovedisplayshortskip}{0pt}
\setlength{\belowdisplayshortskip}{2pt}




\ifllncs
\ifanonymous
\author{}
\institute{}
\title{\titletext}
\else
\title{\titletext}
\author{Mihir Bellare\inst{1} \and Hannah Davis\inst{1} \and Zijing Di\inst{2}}
\institute{University of California San Diego \\ \email{mihir@eng.ucsd.edu}, \email{h3davis@eng.ucsd.edu} \and Stanford University}
\fi  % anonymous

\else % not llncs
\ifacm

\else % not acm
\title{\textbf{\titletext\protect\vspace{0.3in}}}
\appearsin{A preliminary version of this paper appears in the Public-Key Cryptography (PKC) 2023 conference. This is the full version.}
\author{\textsc{Mihir Bellare}\thanks{\ \BellareThanks} \and \textsc{Hannah Davis}\thanks{\ \DavisThanks} \and \textsc{Zijing Di}\thanks{\ Stanford University. {\tt zidi@stanford.edu}.} \vspace{0.3in}}
\date{February 2023\vspace{0.2in}}
\fi % acm
\fi % llncs

\maketitle


% \vspace{-0.55in}
\begin{abstract}
\input{abstract}

\ifllncs\ifanonymous\keywords{Standards \and Signatures \and Indifferentiability \and Random-Oracle Model}\fi\fi 
\end{abstract}

\ifacm\else\ifllncs\else
\newpage
\tableofcontents
\newpage
\fi\fi


\input{intro}
\input{prelim}
\input{direct}
%\input{indiff}
%\input{sec-indiff}
\input{schnorr}
\input{chop}

%\input{our-indiff}

%\input{clamp}
%\input{mu}




\section*{Acknowledgments}

\ifllncs Bellare and Davis are supported in part by NSF grant CNS-2154272. \fi We thank the (anonymous) reviewers of Crypto 2022, Asiacrypt 2022 and CT-RSA 2023 for their valuable comments.
We thank Joseph Jaeger for his helpful comments and discussions about the correctness of chop-MD proofs in the literature.




\addcontentsline{toc}{section}{References}
\ifuseshortbib
\input{refs-short}
\else
\begin{small}
\bibliographystyle{abbrv}
\bibliography{../bibfiles/abbrev3,../bibfiles/crypto,main}
\end{small}
\fi

\appendix



% \input{shrink-proof}
%\input{vf}
\input{prop1-proof}
%\input{lem3-pf}
%\input{lem7-pf}
%\input{apx-sample}
\input{gp-instantiation}
%\input{eddsa-md-proof}
%\input{comp-proof}
%\input{f-eddsa-proof}
%\input{md-indiff}
%\input{pvf-svf-proof}
%\input{pub-indiff}


\end{document}
